---
title: "Week 5 Class"
author: "Molly Farry-Thorn"
date: "September 26, 2017"
output: pdf_document
---

```{r}
model.4pi <- lmer(Utterances.with.Letters  ~ 1 + Session + (1 + Session|Subject), data = parents)
summary(model.4pi)
model.4p <- lmer(Utterances.with.Letters  ~ -1 + Session + (-1 + Session|Subject), data = parents)
summary(model.4p)
#log likelihood - larger values better, AIC/BIC smaller is better
anova(model.4pi, model.4p)

model.4ci <- lmer(Utterances.with.Letters  ~ 1 + Session + (1 + Session|Subject), data = children)
summary(model.4ci)
model.4c <- lmer(Utterances.with.Letters  ~ -1 + Session + (-1 + Session|Subject), data = children)
summary(model.4c)
anova(model.4ci, model.4c)
# Go with no intercept


coefs <- data.frame(coef(model.4pi)[[1]])
ggplot(data = parents, aes(Session, Utterances.with.Letters)) +
   stat_smooth(aes(Session, Utterances.with.Letters), method = lm, se = F) +
   xlim(0,12) + ylim (-20,20) +
  geom_abline(data = coefs, aes(slope = Session, intercept = 0), alpha = 0.2)


coefs <- data.frame(coef(model.4c)[[1]])
ggplot(data = children, aes(Session, Utterances.with.Letters)) +
   stat_smooth(aes(Session, Utterances.with.Letters), method = lm, se = F) +
   xlim(0,12) + ylim (-20,20) +
  geom_abline(data = coefs, aes(slope = Session, intercept = 0), alpha = 0.2)
```
from broom package:
augment gives individual
tidy gives fixed effects - gives sds not variances
glace gives fit values

deviance want larger - related to log likelihood
For AIC & BIC we want smaller values

Below is code that should work for all models. Just run the function and save it as an R object. You can use this with papaja and the apa_table() function pretty easily. The trick is that if you are not using the papaja template, the proper LaTeX packages may not be loaded. You can get around this by attaching a .tex file calling the packages under "in_header: header.tex" in your YAML header. The YAML header of this .Rmd file contains the necessary syntax and the header.tex file with the proper packages.

```{r}
fixed %.% filter(term %in% c("Session","Child.Gender"))
library(dplyr)
## here's some code to make a table. You shouldn't need to modify anything here 
# unless you add additional random effects terms
## fixed effects first ##
table_fun <- function(model){
  fixed <- broom::tidy(model) %>% filter(group == "fixed") %>%
    select(term, estimate) 
  ## add random effects ##
  rand <- broom::tidy(model) %>% filter(group != "fixed") %>%
    select(term, estimate)
  ## get confidence intervals ##
  CI <- data.frame(confint.merMod(model, method = "boot", nsim = 10)) %>%
    mutate(term = rownames(.)) %>% setNames(c("lower", "upper", "term"))
  
  ## Get ICC & R2 values ##
  ICC <- reghelper::ICC(model)
  R2 <- MuMIn::r.squaredGLMM(model)
  
  ## format the fixed effects
  fixed <- fixed %>% left_join(CI %>% filter(!grepl(".sig", term))) %>%
    mutate(type = "Fixed Parts")
  
  rand <- rand %>%
    mutate(estimate = ifelse(grepl("cor", term) == T, estimate, estimate^2),
          term = mapvalues(term, unique(term), 
           c("$\\tau_{00}$", "$\\tau_{11}$", "$\\tau_{10}$", "$\\hat{\\sigma^2}$"))) %>%
    left_join(
      CI %>% filter(grepl(".sig", term)) %>%
        mutate(term = mapvalues(term, unique(term), 
            c("$\\tau_{00}$", "$\\tau_{10}$", "$\\tau_{11}$", "$\\hat{\\sigma^2}$")),
            lower = lower^2, upper = upper^2)) %>%
    mutate(type = "Random Parts")
  
  mod_terms <- tribble(
    ~term, ~estimate, ~type,
    "ICC", ICC, "Model Terms",
    "$R^2_m$", R2[1], "Model Terms",
    "$R^2_c$", R2[2], "Model Terms"
  )
  
  tab <- fixed %>%
    full_join(rand) %>%
    mutate(CI = sprintf("(%.2f, %.2f)", lower, upper)) %>%
    select(-lower, -upper) %>%
    full_join(mod_terms) %>%
    mutate(estimate = sprintf("%.2f", estimate)) %>%
    dplyr::rename(b = estimate) %>%
    select(type, everything())
  return(tab)
}
# you can use this with papaja and the apa_table function pretty easily
# the trick is that if you are not using the papaja template, the proper
# LaTeX packages may not be loaded. You can get around this by attaching
# a .tex file calling the packages under "in_header: header.tex" in your YAML
# header the YAML header of this .Rmd file contains the necessary syntax and 
# the header.tex file with the proper packages

tab <- table_fun(mod2g)

# copy latex into new markdown file and select output Word
output:
  word_document:
    includes:
      in_header:
        header.tex

```

```{r, results= 'asis'}
library(kableExtra)
options(knitr.kable.NA = '')
knitr::kable(tab %>% #select(-type) %>%
    mutate(term = gsub("[()]", "", term)),
             caption = "Not Quite Right kableExtra MLM Table Example", 
    format = "latex", 
    #longtable = T, 
    booktabs = T, escape = F) %>%
  # group_rows("Fixed", 1,4) %>% 
  # group_rows("Random", 5,9) %>%
  # group_rows("Model", 9,11) %>%
  collapse_rows(1) %>%
  #kable_styling(latex_options = c("striped","repeat_header"),full_width = F)
  add_header_above(c(" ", " ", "Model 1" = 2))


# apa_table has col_spanners - which columns to span and what to name it 
# col_spanners = list('Sensation Seeking' = c(2,3))
# col.names = c(",", rep(c("b,"CI)), times = 3)
# align = c("l","l",)

papaja::apa_table(tab %>% select(-type),caption = "papaja MLM Table Example", 
    na_string = "", stub_indents = list(Fixed = c(1:4), Random = c(5:11), Summary = c(9:11)))
```

```{r}
library(sjPlot)
# Categorical, to show the interaction
sjp.int(mod2g, type = "eff", p.kr = F, swap.pred = T)

# Continuous
sjp.int(mod1b, type = "eff", p.kr = F, swap.pred = T, mdrt.values = "meansd")

```

```{r}
# Plotting with ggplot
# example for categorical
fixed.frame <- 
  data.frame(
    expand.grid(
      # here, you add values for your time variable and predictors, for me: graph from 1 to 12 , how often doesn't really matter
      # Want a column for each of the terms in your model
      age0 = seq(0,10,2), 
      groupsNone = c(0,1))) %>%
  # now take care of interactions and add an intercept
  mutate(`age0:groupsNone` = age0*groupsNone,
         Intercept = 1) %>%
  # reordering everything
  select(Intercept, everything())
```

```{r}
# Categorical again - nice for more complex models, don't have to individually deal with interactions
# Using predict to use age and group and plugging into an equation
fixed.frame <- data.frame(expand.grid(age0 = seq(0,10,2), 
      groups = c("None","Jail"))) %>% # call them by name
  mutate(pred = predict(mod2g, newdata = ., re.form = NA)) # not using random effects

fixed.frame$pred <- as.vector(as.matrix(fixed.frame) %*% fixef(mod2g))

fixed.frame %>%
  mutate(age = age0 + 14) %>%
  ggplot(aes(x = age, y = pred, color = groups)) +
    geom_line(size = 2) + 
    lims(y = c(0,4)) +
    labs(x = "Age", y = "Sensation Seeking Composite",
         title = "2 Group Time Invariant Conditional Growth Models") +
    theme_classic() +
    theme(axis.text = element_text(face = "bold", size = rel(1.2)),
          axis.title = element_text(face = "bold", size = rel(1.2)),
          legend.title = element_text(face = "bold", size = rel(1.2)),
          plot.title = element_text(face = "bold", size = rel(1.2), hjust = .5))
```

```{r}
# multiplying to get values for model frame
fixed.frame$pred <- as.vector(as.matrix(fixed.frame) %*% fixef(mod2g))

fixed.frame %>%
  mutate(groups = factor(groupsNone, levels = c(0,1), labels = c("Jail", "None")),
         age = age0 + 14) %>%
  ggplot(aes(x = age, y = pred, color = groups)) +
    geom_line(size = 2) + 
    labs(x = "Age", y = "Sensation Seeking Composite",
         title = "2 Group Time Invariant Conditional Growth Models") +
    theme_classic() +
    theme(axis.text = element_text(face = "bold", size = rel(1.2)),
          axis.title = element_text(face = "bold", size = rel(1.2)),
          legend.title = element_text(face = "bold", size = rel(1.2)),
          plot.title = element_text(face = "bold", size = rel(1.2), hjust = .5))
```

```{r}
# example for continuous
fixed.frame <- sample_dat %>%
  summarise(mean = mean(DemPweight, na.rm = T), 
            sd = sd(DemPweight, na.rm = T))

fixed.frame <- 
  data.frame(
    expand.grid(
      # here, you add values for your time variable and predictors
      age0 = seq(0,10,2), 
      DemPweight = c(fixed.frame$mean-fixed.frame$sd,
                     fixed.frame$mean,
                     fixed.frame$mean+fixed.frame$sd))) %>%
    mutate(pred = predict(mod1b, newdata = ., re.form =NA))
  # now take care of interactions and add an intercept
  #mutate(`age0:DemPweight` = age0*DemPweight,
         #Intercept = 1) %>%
  # reordering everything
  #select(Intercept, everything())

# multiplying to get values for model frame
#fixed.frame$value <- as.vector(as.matrix(fixed.frame) %*% fixef(mod1b))


fixed.frame %>%
  mutate(Weight = factor(DemPweight, levels = unique(DemPweight), labels = c("-1SD", "0SD", "1SD")),
         age = age0 + 14) %>%
  ggplot(aes(x = age, y = value, color = Weight)) +
    geom_line(size = 2) + 
    labs(x = "Age", y = "Sensation Seeking Composite",
         title = "Continuous Invariant Conditional Growth Models") +
    theme_classic() +
    theme(axis.text = element_text(face = "bold", size = rel(1.2)),
          axis.title = element_text(face = "bold", size = rel(1.2)),
          legend.title = element_text(face = "bold", size = rel(1.2)),
          plot.title = element_text(face = "bold", size = rel(1.2), hjust = .5))
```

γ00 - intercept for group coded as 0
γ01 - difference between the two groups
Why do people vary on the intercept? Now we're saying group might explain it. Random effects (μ0j) will get smaller if group is explaining something. 

If group is also predicting slope. Why do some people have shallower slopes? Does group help explain that. μ1j should get smaller. Will also change the γ10 - it's serving as an intercept - it's the average slope and we're adjusting up and down based on someone's intercept and residual.
Any time there is a level 2 predictor of slope you get an interaction. Any time there is a predictor of slope, you have an interaction.  
No random effect (got rid of μ1j) - then you assume that within groups they have the same slopes but group slopes can be different. Might not need a random slope once you have added a group - group might explain the randomness.

Maybe I should change the unit of time...
Rescale my proportion?

Controlling for job status - shifting the slope. The slope is interpreted as accounting for Job. What would happen if everyone had same effect of job.
Job coefficient - how big a hit do you take if you don't have a job, controlling for time.
