---
title: "Week 9 Homework"
author: "Molly Farry-Thorn"
date: "November 2, 2017"
output: pdf_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
library(psych)
library(lavaan)
library(semPlot)
library(semTools)
library(tidyr)
library(plyr)

# Set up my data
Chicago <- read.delim("12.15 CTOPP & WJ scores.csv", header=TRUE, sep=",")
Chicago.wide <- gather(Chicago, Var, Val, select=c("Age", "Age.c", "Timepoint", "ctopp_el_raw", "ctopp_nr_raw", "ctopp_sm_raw", "wj_wordid_raw", "wj_wordattack_raw","wj_wordid_standard", "Sex")) %>%
           unite(VarG, Var, Grade) %>%
           spread(VarG, Val)
names(Chicago.wide)[1] <- "Participant"
Chicago.wide <- Chicago.wide[,c(1,8:9,6:7,4:5,2:3,24:25,22:23,28:29,26:27,
                                12,16,20,32,36,40,
                                13,17,21,33,37,41,
                                10,14,18,30,34,38,
                                11,15,19,31,35,39)]
names(Chicago)[1] <- "Participant"

```

# 1. Fit a measurement model to your constructs at one time point. Try out the different types of scaling discussed in class. What changes what stays the same?

```{r, warning = FALSE}
mod.1 <- 'phonemic =~ ctopp_el_raw_KG_1 + ctopp_nr_raw_KG_1 + ctopp_sm_raw_KG_1
          reading =~ wj_wordid_raw_KG_1 + wj_wordattack_raw_KG_1'

fit.1 <- cfa(mod.1, data=Chicago.wide)

summary(fit.1, fit.measures=TRUE, standardized = TRUE) # Josh included standardized=TRUE. What does that mean?

semPaths(fit.1, layout = "tree", whatLabels = "est")
# semPaths(fit.1, layout = "tree", what = "std") What is this doing??
# no triangles means no mean structure - which is fine when not doing longitudinal (think of those triangles as zero)

# Fixed factor approach rather than a marker variable approach

mod.2 <- 'phonemic =~ ctopp_el_raw_KG_1 + ctopp_nr_raw_KG_1 + ctopp_sm_raw_KG_1
          reading =~ wj_wordid_raw_KG_1 + wj_wordattack_raw_KG_1'

fit.2 <- cfa(mod.2, std.lv=TRUE, data=Chicago.wide)

summary(fit.2, fit.measures=TRUE,  standardized = TRUE)

semPaths(fit.2, layout = "tree", whatLabels = "est")
```
Under the marker method, parameter estimates (means) were fixed at 1 for the first indicators (i.e. ctopp_el_raw_KG_1 and wj_wordid_raw_KG_1). 

Under the fixed factor method, in contrast, parameter estimates (variances) were fixed to 1 for the latent variables (i.e. phonemic and reading). 

In both cases, fit indices (e.g., logLikelihood, TLI, CFI, RMSEA) remain constant.

# 2. What do the fit statistics say about your latent variable? Good/bad? Is your latent variable Just identified/saturdated, under identified or over identified?

RMSEA = .000, SRMR = .035, TLI = 1.02, CFI = 1.00

TLI and CFI > .90 and RMSEA and SRMR < .08, all suggesting good fit.

This model is over identified, as evidenced by the positive degrees of freedom (4).

# 3. Fit a longitudinal CFA model where you a) first correlate your latent factors across time and then b) a second model that predicts later times by a prevous time (ie auto regressive; t1 -> t2 -> t3). What are your conclusions? How does one differ from the other?

```{r, warning = FALSE}
# latent variables correlated across time
mod.3 <- '
## define latent variables
          Phonemic_t1 =~ ctopp_el_raw_KG_1 + ctopp_nr_raw_KG_1 + ctopp_sm_raw_KG_1 
          Phonemic_t2 =~ ctopp_el_raw_KG_3 + ctopp_nr_raw_KG_3 + ctopp_sm_raw_KG_3
          Phonemic_t3 =~ ctopp_el_raw_FG_1 + ctopp_nr_raw_FG_1 + ctopp_sm_raw_FG_1

## correlated residuals across time
          ctopp_el_raw_KG_1 ~~ ctopp_el_raw_KG_3 + ctopp_el_raw_FG_1
          ctopp_el_raw_KG_3 ~~ ctopp_el_raw_FG_1
          ctopp_nr_raw_KG_1 ~~ ctopp_nr_raw_KG_3 + ctopp_nr_raw_FG_1
          ctopp_nr_raw_KG_3 ~~ ctopp_nr_raw_FG_1
          ctopp_sm_raw_KG_1 ~~ ctopp_sm_raw_KG_3 + ctopp_sm_raw_FG_1
          ctopp_sm_raw_KG_3 ~~ ctopp_sm_raw_FG_1'

fit.3 <- cfa(mod.3,data = Chicago.wide, meanstructure=TRUE, std.lv = TRUE)
summary(fit.3, standardized=TRUE, fit.measures=TRUE)
semPaths(fit.3, whatLabels = "std")

#Auto regressive, later times predicted by earlier times
mod.4 <- 'Phonemic_t1 =~ L1*ctopp_el_raw_KG_1 + L2*ctopp_nr_raw_KG_1 + L3*ctopp_sm_raw_KG_1 
          Phonemic_t2 =~ L1*ctopp_el_raw_KG_3 + L2*ctopp_nr_raw_KG_3 + L3*ctopp_sm_raw_KG_3
          Phonemic_t3 =~ L1*ctopp_el_raw_FG_1 + L2*ctopp_nr_raw_FG_1 + L3*ctopp_sm_raw_FG_1

##correlated residuals across time
          ctopp_el_raw_KG_1 ~~ ctopp_el_raw_KG_3 + ctopp_el_raw_FG_1
          ctopp_el_raw_KG_3 ~~ ctopp_el_raw_FG_1
          ctopp_nr_raw_KG_1 ~~ ctopp_nr_raw_KG_3 + ctopp_nr_raw_FG_1
          ctopp_nr_raw_KG_3 ~~ ctopp_nr_raw_FG_1
          ctopp_sm_raw_KG_1 ~~ ctopp_sm_raw_KG_3 + ctopp_sm_raw_FG_1
          ctopp_sm_raw_KG_3 ~~ ctopp_sm_raw_FG_1

##directional regression paths
          Phonemic_t3 ~ Phonemic_t2
          Phonemic_t2 ~ Phonemic_t1

## free latent variances at later times (only set the scale once)
          Phonemic_t2 ~~ NA*Phonemic_t2
          Phonemic_t3 ~~ NA*Phonemic_t3'

fit.4 <- sem(mod.4, data=Chicago.wide, std.lv = TRUE)
summary(fit.4, fit.measures = TRUE)
semPaths(fit.4, whatLabels = "std")
```
For the longitudinal CFA model with correlated latent factors (mod.3), I conclude that my latent factors are
strongly correlated across time. Moreover, across all three time points, indicators load strongly & significantly
onto their respective latent constructs.
For the autoregressive CFA model (mod.4), a similar picture emerges, suggesting that Semantic_1 is
highly predictive of Semantic_2, which is highly predictive of Semantic_3. Of note, standardized variances
for S_2 and S_3 are insignificant (p = .10 and .62, respectively), likely due to the fact that their variability
is accounted for by

# 4. Fit a longitdinal growth model in SEM and in HLM. Compare and contrast the differences.
```{r, warning = FALSE}
library(lme4)

mod.HLM <- lmer(wj_wordid_raw ~ Timepoint + (1 | Participant), data = Chicago) #fixed slope
summary(mod.HLM) # intercept: 17.53, slope: 7.24

mod.HLM2 <- lmer(wj_wordid_raw ~ Timepoint + (Timepoint | Participant), data = Chicago) #fixed slope
summary(mod.HLM2) # intercept: 17.51, slope: 7.24

mod.SEM <- 'intercept =~ 1*wj_wordid_raw_KG_1 + 1*wj_wordid_raw_KG_3 + 1*wj_wordid_raw_FG_1 + 1*wj_wordid_raw_FG_3
            slope =~ 0*wj_wordid_raw_KG_1 + 1*wj_wordid_raw_KG_3 + 2*wj_wordid_raw_FG_1 + 3*wj_wordid_raw_FG_3
            slope ~~ 0*slope' #fixed slope, no variance
mod.SEM.fixed <- growth(mod.SEM, missing = "ML", data = Chicago.wide)
summary(mod.SEM.fixed) # intercept: 17.34, slope: 7.11
semPaths(mod.SEM.fixed, what = "std")

mod.SEM2 <- 'intercept =~ 1*wj_wordid_raw_KG_1 + 1*wj_wordid_raw_KG_3 + 1*wj_wordid_raw_FG_1 + 1*wj_wordid_raw_FG_3
             slope =~ 0*wj_wordid_raw_KG_1 + 1*wj_wordid_raw_KG_3 + 2*wj_wordid_raw_FG_1 + 3*wj_wordid_raw_FG_3'
mod.SEM.random <- growth(mod.SEM2, missing = "ML", data = Chicago.wide)
summary(mod.SEM.random) # intercept: 17.28, slope: 7.17
semPaths(mod.SEM.random, what = "std")

```
Estimates of intercept are similar between fixed slope SEM & HLM models and between random slope SEM & HLM models.

Estimates of slope are similar between fixed slope SEM & HLM models and between random slope SEM & HLM models.

# 5. Constrain the residual variances to be equal. Does this change the fit of your model?
Constraining the residual variances does not significantly change model fit. LogLikelihood tests indicate that
a simpler model, where residual variances are allowed to vary, is preferred to a more complex model where
they are constrained to be equal.
```{r}
mod.SEM3 <- 'intercept =~ 1*wj_wordid_raw_KG_1 + 1*wj_wordid_raw_KG_3 + 1*wj_wordid_raw_FG_1 + 1*wj_wordid_raw_FG_3
             slope =~ 0*wj_wordid_raw_KG_1 + 1*wj_wordid_raw_KG_3 + 2*wj_wordid_raw_FG_1 + 3*wj_wordid_raw_FG_3

wj_wordid_raw_KG_1 ~~ a*wj_wordid_raw_KG_1
wj_wordid_raw_KG_3 ~~ a*wj_wordid_raw_KG_3
wj_wordid_raw_FG_1 ~~ a*wj_wordid_raw_FG_1
wj_wordid_raw_FG_3 ~~ a*wj_wordid_raw_FG_3' #random slope, residual variances equal

mod.SEM.equal.resid <- growth(mod.SEM3, missing = "ML", data = Chicago.wide)
summary(mod.SEM.equal.resid) #intercept = 17.51, slope = 7.24
semPaths(mod.SEM.equal.resid)
anova(mod.SEM.random, mod.SEM.equal.resid) 
```
Constraining the residual variances does not significantly change model fit. LogLikelihood tests indicate that
a simpler model, where residual variances are allowed to vary, is preferred to a more complex model where
they are constrained to be equal.

# 6. Contrain your slope to be fixed, not random. How does this change your model?
```{r}
summary(mod.SEM.fixed) # model from question #4 with fixed slope.

anova(mod.SEM.random, mod.SEM.fixed) 
```


Constraining slopes to be fixed does not significantly change my model (p = .78).

# 7. Change the time metric in your SEM growth model. How does that change your estimates? Does it change your fit statistics?
changed my time metric such that the intercept was centered at the end of first grade rather than at the beginning of kindergarten. This increased the intercept from around 17.5 to 38.8 (which makes sense, given age-related change) but the slope did not change much and model fit did not change.

```{r}
mod.SEM4 <- 'intercept =~ 1*wj_wordid_raw_KG_1 + 1*wj_wordid_raw_KG_3 + 1*wj_wordid_raw_FG_1 + 1*wj_wordid_raw_FG_3
             slope =~ -3*wj_wordid_raw_KG_1 + -2*wj_wordid_raw_KG_3 + -1*wj_wordid_raw_FG_1 + 0*wj_wordid_raw_FG_3'
mod.SEM.time <- growth(mod.SEM4, missing = "ML", data = Chicago.wide)
summary(mod.SEM.time)

anova(mod.SEM.random, mod.SEM.time)

```


# 8. Try a different type of estimation (see lavaan tutorial for details). How does that change your model?
Changing the estimator to MLR did not change the estimates. Instead, it changes the standard errors. And it added an extra "Robust" column to the fit statistics to indicate the robust goodness of fit.
```{r}
mod.SEM.random.MLR <- growth(mod.SEM2, missing = "ML", estimator = "MLR", data = Chicago.wide)
summary(mod.SEM.random)
summary(mod.SEM.random.MLR)

```

#—-NEW QUESTIONS—-

# 10. Test measurement invariance across time for your construct. Can you run growth models? If there is evidence of non-invariance, what seems to be the problem?
The config model fits significantly better than the weak model and the cfi decreases by more than .01, suggesting measurement invariance.
```{r, warning = FALSE}

config.fit <- fit.3

summary(config.fit, standardized = TRUE, fit.measures = T)
semPaths(config.fit)

# Weak - constrain factor loadings to be the same across time

weak <- '
## define latent variables, constrain factor loadings
          Phonemic_t1 =~ L1*ctopp_el_raw_KG_1 + L2*ctopp_nr_raw_KG_1 + L3*ctopp_sm_raw_KG_1 
          Phonemic_t2 =~ L1*ctopp_el_raw_KG_3 + L2*ctopp_nr_raw_KG_3 + L3*ctopp_sm_raw_KG_3
          Phonemic_t3 =~ L1*ctopp_el_raw_FG_1 + L2*ctopp_nr_raw_FG_1 + L3*ctopp_sm_raw_FG_1

## free latent variances at later times
Phonemic_t2 ~~ NA*Phonemic_t2
Phonemic_t3 ~~ NA*Phonemic_t3

## correlated residuals across time
          ctopp_el_raw_KG_1 ~~ ctopp_el_raw_KG_3 + ctopp_el_raw_FG_1
          ctopp_el_raw_KG_3 ~~ ctopp_el_raw_FG_1
          ctopp_nr_raw_KG_1 ~~ ctopp_nr_raw_KG_3 + ctopp_nr_raw_FG_1
          ctopp_nr_raw_KG_3 ~~ ctopp_nr_raw_FG_1
          ctopp_sm_raw_KG_1 ~~ ctopp_sm_raw_KG_3 + ctopp_sm_raw_FG_1
          ctopp_sm_raw_KG_3 ~~ ctopp_sm_raw_FG_1'

weak.fit <- cfa(weak, data=Chicago.wide, meanstructure=TRUE, std.lv=TRUE)

summary(weak.fit, standardized=TRUE, fit.measures=TRUE)
semPaths(weak.fit)

# Strong - constrain means/intercepts of indicators to be equal

strong <- '
## define latent variables, constrain factor loadings
          Phonemic_t1 =~ L1*ctopp_el_raw_KG_1 + L2*ctopp_nr_raw_KG_1 + L3*ctopp_sm_raw_KG_1 
          Phonemic_t2 =~ L1*ctopp_el_raw_KG_3 + L2*ctopp_nr_raw_KG_3 + L3*ctopp_sm_raw_KG_3
          Phonemic_t3 =~ L1*ctopp_el_raw_FG_1 + L2*ctopp_nr_raw_FG_1 + L3*ctopp_sm_raw_FG_1

## free latent variances at later times
Phonemic_t2 ~~ NA*Phonemic_t2
Phonemic_t3 ~~ NA*Phonemic_t3


## correlated residuals across time
          ctopp_el_raw_KG_1 ~~ ctopp_el_raw_KG_3 + ctopp_el_raw_FG_1
          ctopp_el_raw_KG_3 ~~ ctopp_el_raw_FG_1
          ctopp_nr_raw_KG_1 ~~ ctopp_nr_raw_KG_3 + ctopp_nr_raw_FG_1
          ctopp_nr_raw_KG_3 ~~ ctopp_nr_raw_FG_1
          ctopp_sm_raw_KG_1 ~~ ctopp_sm_raw_KG_3 + ctopp_sm_raw_FG_1
          ctopp_sm_raw_KG_3 ~~ ctopp_sm_raw_FG_1

## constrain intercepts across time
ctopp_el_raw_KG_1 ~ A*1
ctopp_nr_raw_KG_1 ~ B*1
ctopp_sm_raw_KG_1 ~ C*1

ctopp_el_raw_KG_3 ~ A*1
ctopp_nr_raw_KG_3 ~ B*1
ctopp_sm_raw_KG_3 ~ C*1

ctopp_el_raw_FG_1 ~ A*1
ctopp_nr_raw_FG_1 ~ B*1
ctopp_sm_raw_FG_1 ~ C*1

##free latent means at later times
          Phonemic_t2 ~ NA*1
          Phonemic_t3 ~ NA*1'

strong.fit <- cfa(strong, data=Chicago.wide, meanstructure=TRUE, std.lv=TRUE)

summary(strong.fit, standardized=TRUE, fit.measures=TRUE)

semPaths(strong.fit)

anova(config.fit, weak.fit)
anova(weak.fit, strong.fit)

fitmeasures(config.fit)['cfi']
fitmeasures(weak.fit)['cfi']
fitmeasures(strong.fit)['cfi']

```


# 11. Fit a second order growth model. Compare and contrast the estimates with the normal latent growth model.
Overall the estimates are not that different. The estimates of the first order model vs. the second order model:

Intercept: 6.734 vs. 7.445
Slope: 1.426 vs. 1.696
```{r, warning = FALSE}
sec.order <- '
## define latent variables
          Phonemic_t1 =~ NA*L1*ctopp_el_raw_KG_1 + L1*ctopp_el_raw_KG_1 + L2*ctopp_nr_raw_KG_1 + L3*ctopp_sm_raw_KG_1 
          Phonemic_t2 =~ NA*L1*ctopp_el_raw_KG_3 + L1*ctopp_el_raw_KG_3 + L2*ctopp_nr_raw_KG_3 + L3*ctopp_sm_raw_KG_3
          Phonemic_t3 =~ NA*L1*ctopp_el_raw_FG_1 + L1*ctopp_el_raw_FG_1 + L2*ctopp_nr_raw_FG_1 + L3*ctopp_sm_raw_FG_1


## intercepts
ctopp_el_raw_KG_1 ~ A*1
ctopp_nr_raw_KG_1 ~ B*1
ctopp_sm_raw_KG_1 ~ C*1

ctopp_el_raw_KG_3 ~ A*1
ctopp_nr_raw_KG_3 ~ B*1
ctopp_sm_raw_KG_3 ~ C*1

ctopp_el_raw_FG_1 ~ A*1
ctopp_nr_raw_FG_1 ~ B*1
ctopp_sm_raw_FG_1 ~ C*1


## correlated residuals across time
          ctopp_el_raw_KG_1 ~~ ctopp_el_raw_KG_3 + ctopp_el_raw_FG_1
          ctopp_el_raw_KG_3 ~~ ctopp_el_raw_FG_1
          ctopp_nr_raw_KG_1 ~~ ctopp_nr_raw_KG_3 + ctopp_nr_raw_FG_1
          ctopp_nr_raw_KG_3 ~~ ctopp_nr_raw_FG_1
          ctopp_sm_raw_KG_1 ~~ ctopp_sm_raw_KG_3 + ctopp_sm_raw_FG_1
          ctopp_sm_raw_KG_3 ~~ ctopp_sm_raw_FG_1


## latent variable intercepts
Phonemic_t1  ~ 0*1
Phonemic_t2  ~ 0*1
Phonemic_t3  ~ 0*1

#model constraints for effect coding
## loadings must average to 1
L1 == 3 - L2 - L3
## means must average to 0
A == 0 - B - C

i =~ 1*Phonemic_t1 + 1*Phonemic_t2 + 1*Phonemic_t3 
s =~ 0*Phonemic_t1 + 1*Phonemic_t2 + 2*Phonemic_t3 '

fit.sec.order <- growth(sec.order, data=Chicago.wide, meanstructure = TRUE, missing = "ML")

summary(fit.sec.order, fit.measures=TRUE)

#Normal latent growth model
first.order <- '
              i =~ 1*ctopp_el_raw_KG_1 + 1*ctopp_nr_raw_KG_1 + 1*ctopp_sm_raw_KG_1 +
                   1*ctopp_el_raw_KG_3 + 1*ctopp_nr_raw_KG_3 + 1*ctopp_sm_raw_KG_3 +
                   1*ctopp_el_raw_FG_1 + 1*ctopp_nr_raw_FG_1 + 1*ctopp_sm_raw_FG_1
              s =~ 0*ctopp_el_raw_KG_1 + 0*ctopp_nr_raw_KG_1 + 0*ctopp_sm_raw_KG_1 +
                   1*ctopp_el_raw_KG_3 + 1*ctopp_nr_raw_KG_3 + 1*ctopp_sm_raw_KG_3 +
                   2*ctopp_el_raw_FG_1 + 2*ctopp_nr_raw_FG_1 + 2*ctopp_sm_raw_FG_1'

fit.first.order <- growth(first.order, data=Chicago.wide, meanstructure = TRUE, missing = "ML")
summary(fit.first.order, fit.measures=TRUE)

```


# 12. Fit a series of multiple group models. Constrain some parameters and compare the fit.
```{r}
group1 <- '
## define latent variables
          Phonemic_t1 =~ ctopp_el_raw_KG_1 + ctopp_nr_raw_KG_1 + ctopp_sm_raw_KG_1 
          Phonemic_t2 =~ ctopp_el_raw_KG_3 + ctopp_nr_raw_KG_3 + ctopp_sm_raw_KG_3
          Phonemic_t3 =~ ctopp_el_raw_FG_1 + ctopp_nr_raw_FG_1 + ctopp_sm_raw_FG_1

## correlated residuals across time
          ctopp_el_raw_KG_1 ~~ ctopp_el_raw_KG_3 + ctopp_el_raw_FG_1
          ctopp_el_raw_KG_3 ~~ ctopp_el_raw_FG_1
          ctopp_nr_raw_KG_1 ~~ ctopp_nr_raw_KG_3 + ctopp_nr_raw_FG_1
          ctopp_nr_raw_KG_3 ~~ ctopp_nr_raw_FG_1
          ctopp_sm_raw_KG_1 ~~ ctopp_sm_raw_KG_3 + ctopp_sm_raw_FG_1
          ctopp_sm_raw_KG_3 ~~ ctopp_sm_raw_FG_1

Phonemic_t1 ~~ NA*Phonemic_t2 + NA*Phonemic_t3
Phonemic_t2 ~~ c(1,1)*Phonemic_t3'


fit.group.1 <- cfa(group1, data=Chicago.wide, std.lv=TRUE, group = "Sex_KG_1")

summary(fit.group.1, standardized=TRUE, fit.measures=TRUE)

fit.group.2 <- cfa(group1, data=Chicago.wide, std.lv=TRUE, group = "Sex_KG_1", group.equal = c("loadings", "intercepts"))
summary(fit.group.2)
anova(fit.group.1, fit.group.2)
```

