---
title: "Week 5 Homework"
author: "Molly Farry-Thorn"
date: "October 3, 2017"
output: pdf_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
library(psych)
library(dplyr)
library(sjPlot)
library(tidyverse)
library(broom)
library(lme4)
library(MuMIn)
library(merTools)
library(reghelper)
library(stargazer)
library(lsmeans)
library(multcompView)
library(ggplot2)
letters <- read.delim("Masters coded letters.csv", header=TRUE, sep=",")
letters$Prop.Letters <- letters$Utterances.with.Letters/letters$Utterances
letters$Subject <- as.factor(letters$Subject)
letters$Speaker <- as.factor(letters$Speaker)
letters$Child.Gender <- as.factor(letters$Child.Gender)
levels(letters$Child.Gender) <- c("Male","Female")
parents <- subset(letters, Speaker==0)
children <- subset(letters, Speaker==1)
```

# 1. Run a series of models using a time-invariant nominal covariate. For all models, how does your model change from model to model. What is your final model?

# a) where the covariate only predicts the intercept 
```{r}
# time invariant covariate that predicts the intecept but not slope
children$Child.Gender <- relevel(children$Child.Gender, ref = "Male")

mod1a <- lmer(Utterances.with.Letters ~ Time + Child.Gender + (1|Subject), data = children)
summary(mod1a)
```
**Fixed Effects** 

Intercept: 1.8813; mean number of utterances with letters for male children at Session 1

Time: 2.63; increase in number of utterances with letters every year

Child.Gender: -1.95; difference between males and females at Session 1, females start lower

# b) predicts both intercept and slope 
```{r}
# time invariant predictor for the intercept AND slopes
mod1b <- lmer(Utterances.with.Letters ~ Time + Child.Gender + Time*Child.Gender + (1|Subject), data = children)
summary(mod1b)
```
**Fixed Effects** 

Intercept: 0.29; mean number of utterances with letters for male children at Session 1

Time: 3.50; increase in number of utterances with letters every year, for males

Child.Gender: 1.40; difference between males and females at Session 1, females start higher (different than last model)

Time:Child.Gender: -1.83; difference in the slopes between males and females, the effect of age is smaller in females

# c) is rescaled (e.g. centering). 
```{r}
# changing dummy coding such that reference group becomes females
children$Child.Gender <- relevel(children$Child.Gender, ref = "Female")
mod1c <- lmer(Utterances.with.Letters ~ Time + Child.Gender + Time*Child.Gender + (1|Subject), data = children)
summary(mod1c)
```
**Fixed Effects** 

Intercept: 1.69; mean number of utterances with letters for female children at Session 1

Time: 1.67; increase in number of utterances with letters every year, for females

Child.Gender: -1.40; difference between males and females at Session 1, males start lower. This is the same difference we found in mod1b, just in the opposite direction.

Time:Child.Gender: 1.83; difference in the slopes between males and females, the effect of age is larger in males. This is the same difference we found in mod1b, just in the opposite direction.

```{r}
anova(mod1a,mod1b)
```

The likelihood ratio test suggests that simpler model, where covariate only predicts intercept, is preferred. 

# 2. Introduce a time-invariant continuous covariate and run models a-c from #1.
```{r}
# time-invariant continuous covariate that only predicts the intercept
mod2a <- lmer(Utterances.with.Letters ~ Time + SES + (1|Subject), data = children)
summary(mod2a)
```
**Fixed Effects** 

Intercept: 0.96; mean of number of utterances with letters for children with mean level SES at Session 1

Time: 2.63; increase in number of utterances with letters every year, when SES at mean level 

SES: 1.20; increase in number of utterances for every 1 unit increase in SES

```{r}
# time-invariant continuous covariate predicts the intercept AND slopes
mod2b <- lmer(Utterances.with.Letters ~ Time + SES + Time*SES + (1|Subject), data = children)
summary(mod2b)
```
**Fixed Effects** 

Intercept: 0.96; mean of number of utterances with letters for children with mean level SES at Session 1

Time: 2.63; increase in number of utterances with letters every year, when SES at mean level 

SES: 0.01; increase in number of utterances for every 1 unit increase in SES at Session 1

Time:SES: 0.65; the change in the relationship in the time slope for every 1 unit increase in SES 

*My SES variable is already centered.*
```{r}
anova(mod2a,mod2b)
```
Again, the likelihood ratio test suggests that simpler model, where covariate only predicts intercept, is preferred. 

# 3. Graph both of your final models for the continuous and nominal models above. 
```{r}
# Graphing nominal, where gender only predict the intercept
children$Child.Gender <- relevel(children$Child.Gender, ref = "Male")
sjp.lmer(mod1a, type = "pred.fe", var = c("Time", "Child.Gender"), facet = FALSE, show.scatter = FALSE)

fixed.frame <- data.frame(expand.grid(Time = seq(0,4,1), Child.Gender = c("Male", "Female"))) %>%
  mutate(pred = predict(mod1a, newdata = ., re.form = NA))

nom <- ggplot(aes(x = Time, y = pred, color = Child.Gender), data = fixed.frame) +
  geom_line() +
  labs(x = "Time", y = "# of Utterances with Letters") +
  theme_classic()
nom
```

```{r}
# Graphing continuous, where SES only predicts the intercept
fixed.frame2 <- letters %>%
  summarise(mean = mean(SES, na.rm = T), sd = sd(SES, na.rm = T))
fixed.frame2 <- data.frame(
  expand.grid(
    Time = seq(0,4,1),
    SES = c(fixed.frame2$mean - fixed.frame2$sd,
            fixed.frame2$mean,
            fixed.frame2$mean + fixed.frame2$sd))) %>%
  mutate(pred = predict(mod2a, newdata = ., re.form = NA))

fixed.frame2$SES <- as.factor(fixed.frame2$SES)
levels(fixed.frame2$SES) <- c("-1 SD","Mean", "+1 SD")

con <- ggplot(aes(x = Time, y = pred, color = SES), data = fixed.frame2) +
  geom_line() +
  labs(x = "Time", y = "# of Utterances with Letters") +
  theme_classic()
con
```


# 4. Calculate confidence intervals around your estimates for your final models
```{r}
#Confidence intervals around nominal model
mod1a.ci <- confint(mod1a, level = .95, oldNames = F, method = "boot", nsim = 1000)
broom::tidy(mod1a.ci)

#Confidence intervals around continuous model
mod2a.ci <- confint(mod2a, level = .95, oldNames = F, method = "boot", nsim = 1000)
broom::tidy(mod2a.ci)
```

# 5. Include both types of covariates in a single model. How does your interpretation of parameters change?
```{r}
mod3 <- lmer(Utterances.with.Letters ~ Time*Child.Gender*SES + (1|Subject), data = children)
summary(mod3)
```
**Fixed Effects** 

Intercept: 0.47; mean of number of utterances with letters for males with mean level SES at Session 1

Time: 3.21; increase in number of utterances with letters every year, for males and when SES at mean level 

Child.Gender: 1.35; the difference in the number of utterances with letters between males and females when SES at mean level

SES: -1.25; decrease, for males, in number of utterances for every 1 unit increase in SES at Session 1

Session:Child.Gender: -1.59; the difference between the slopes for males and females when SES at mean level

Time:SES: 2.03; the change in the slope of Time, for males, for every 1 unit increase in SES 

Child.Gender:SES: 2.14; the difference between the relationship of SES and the number of utterances with letters for males and females at Session 1

Time:Child.Gender:SES: -2.37; the difference between the interaction of SES and Time for males and females

# 6. If you have one available, introduce a time-varying covariate.
```{r}
mod4 <- lmer(Utterances.with.Letters ~ Time + Utterances + (1|Subject), data = children)
summary(mod4)
```
**Fixed Effects** 

Intercept: -0.19; mean of number of utterances with letters for children at Session 1

Time: 0.58; increase in number of utterances with letters every year 

Utterances: 0.0033; increase in number of utterances for every 1 unit increase number of utterances