---
title: "Week 6 Class"
author: "Molly Farry-Thorn"
date: "October 5, 2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
library(psych)
library(dplyr)
library(sjPlot)
library(tidyverse)
library(broom)
library(lme4)
library(MuMIn)
library(merTools)
library(reghelper)
library(stargazer)
library(lsmeans)
library(multcompView)
library(ggplot2)
letters <- read.delim("Masters coded letters.csv", header=TRUE, sep=",")
letters$Prop.Letters <- letters$Utterances.with.Letters/letters$Utterances
letters$Subject <- as.factor(letters$Subject)
letters$Speaker <- as.factor(letters$Speaker)
letters$Child.Gender <- as.factor(letters$Child.Gender)
levels(letters$Child.Gender) <- c("Male","Female")
parents <- subset(letters, Speaker==0)
children <- subset(letters, Speaker==1)
```

How should I scale my variables? Should I analyze the proportion?

Sending us daily diary data for next homework

Time is our predictor so we square it. You have two Level 1 predictors - linear and quadratic.
Subtracting out the mean time. So tangent at time 0.
The square is an interaction term. Center it so the lower order terms are easier to interpret.
Without centering - how did that initial trajectory change over time. For really curved trajectories, more helpful to center.
At Level 2 with have average across everyone. We decide if there are random effects or other predictors.
Centering will change your lower order term. Where you're interested in the curve.
Ranef for quadratic - people have different types of curves.
Ranef linear year - people have the same quadratic curve but the linear tangent can be more or less steep.

Splines are helpful when you think there will be two different trajectories before or after a specific event.
Create a knot. Slope up to the knot and slope after the knot.
Difference between initial slope and second slope. Is that difference sig.

# Intensive Data Analysis: Within-Person Fluctuations

Need to insert rows for missing time points
The linear trends aren't the question, instead why are some pople fluctuating
Look at ICC - tells you how much between vs. within. Higher ICC is more between person
Between - some people are just more or less homesick
The mean of personality is better than a mean of mood. 
Level 1 - person center.
Level 2 - grand mean center.

Someone who ruminates more on average might be more homesick - Level 2 

Might not be that ruminating more for yourself makes you more homesick - Level 1

Within person centering - your average is 0 and so you are looking then at your fluctuations and everyone has the same average - deviations from your average. So you're getting rid of the issue that someone who ruminates a lot is more likely to ruminate on a given day.
Aggregate of rumination - mean for each person: This is the between people.
Person centered rumination - variation for each person: This is within pe
