---
title: "Week 3 Homework"
author: "Molly Farry-Thorn"
date: "September 26, 2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
library(tidyverse)
library(lme4)
library(ggplot2)
letters <- read.delim("Masters coded letters.csv", header=TRUE, sep=",")
letters$Subject <- as.factor(letters$Subject)
letters$Speaker <- as.factor(letters$Speaker)
letters$Child.Gender <- as.factor(letters$Child.Gender)
levels(letters$Child.Gender) <- c("Male","Female")
parents <- subset(letters, Speaker==0)
children <- subset(letters, Speaker==1)
```

# 1. Run a series of models using a time-invariant nominal covariate. For all models, how does your model change from model to model. What is your final model?

# a) where the covariate only predicts the intercept 
```{r}
# time invariant covariate that predicts the intecept but not slope
children$Child.Gender <- relevel(children$Child.Gender, ref = "Male")

mod1a <- lmer(Utterances.with.Letters ~ Session + Child.Gender + (Session|Subject), data = children)
summary(mod1a)
```
**Fixed Effects** 

Intercept: -0.30; mean number of utterances with letters for male children at Session 0 (around age 10 months)

Session: 1.17; increase in number of utterances with letters every 4 months

Child.Gender: -0.64; difference between males and females at Session 0, females start lower

# b) predicts both intercept and slope 
```{r}
# time invariant predictor for the intercept AND slopes
mod1b <- lmer(Utterances.with.Letters ~ Session + Child.Gender + Session*Child.Gender + (Session|Subject), data = children)
summary(mod1b)
```
**Fixed Effects** 

Intercept: -0.92; mean number of utterances with letters for male children at Session 0 (around age 10 months)

Session: 1.18; increase in number of utterances with letters every 4 months, for males

Child.Gender: 1.94; difference between males and females at Session 0, females start higher

Session:Child.Gender: -0.60; difference in the slopes between males and females, the effect of age is smaller in females

# c) is rescaled (e.g. centering). 
```{r}
# changing dummy coding such that reference group becomes females
children$Child.Gender <- relevel(children$Child.Gender, ref = "Female")
mod1c <- lmer(Utterances.with.Letters ~ Session + Child.Gender + Session*Child.Gender + (Session|Subject), data = children)
summary(mod1c)
```
**Fixed Effects** 

Intercept: 1.02; mean number of utterances with letters for female children at Session 0 (around age 10 months)

Session: 0.57; increase in number of utterances with letters every 4 months, for females

Child.Gender: -1.94; difference between males and females at Session 0, males start lower. This is the same difference we found in mod1b, just in the opposite direction.

Session:Child.Gender: 0.60; difference in the slopes between males and females, the effect of age is larger. This is the same difference we found in mod1b, just in the opposite direction.

```{r}
anova(mod1a,mod1b)
```

The likelihood ratio test suggests that simpler model, where covariate only predicts intercept, is preferred. 

# 2. Introduce a time-invariant continuous covariate and run models a-c from #1.
```{r}
# time-invariant continuous covariate that only predicts the intercept
mod2a <- lmer(Utterances.with.Letters ~ Session + SES + (Session|Subject), data = children)
summary(mod2a)
```
**Fixed Effects** 

Intercept: 0.00; mean of number of utterances with letters for children with mean level SES at Session 0 (around age 10 months)

Session: 0.89; increase in number of utterances with letters every 4 months, when SES at mean level 

SES: 0.72; increase in number of utterances for every 1 unit increase in SES

```{r}
# time-invariant continuous covariate predicts the intercept AND slopes
mod2b <- lmer(Utterances.with.Letters ~ Session + SES + Session*SES + (Session|Subject), data = children)
summary(mod2b
```
**Fixed Effects** 

Intercept: 0.00; mean of number of utterances with letters for children with mean level SES at Session 0 (around age 10 months)

Session: 0.89; increase in number of utterances with letters every 4 months, when SES at mean level 

SES: -0.19; decrease in number of utterances for every 1 unit increase in SES

Session:SES: 0.21; the change in the slope of Session for every 1 unit increase in SES 

*My SES variable is already centered.*
```{r}
anova(mod2a,mod2b)
```
Again, the likelihood ratio test suggests that simpler model, where covariate only predicts intercept, is preferred. 

# 3. Graph both of your final models for the continuous and nominal models above. 
```{r}
#Graphing nominal
library(sjPlot)
sjp.lmer(mod1a, type = "pred.fe", var = c("Session", "Child.Gender"), facet = FALSE, show.scatter = FALSE)
sjp.int(mod1b, swap.pred = T)

library(broom)

mod1a_aug <- augment(mod1a)
ggplot(mod1a_aug, aes(x= Session, y = .fitted, color=factor(Child.Gender))) + 
  stat_smooth(aes(group=factor(Child.Gender)),method="lm")

mod2a_aug <- augment(mod2a)

ggplot(mod2a_aug, aes(x= Session, y = .fitted)) + 
  geom_point() +
  stat_smooth(aes(group=SES),method="lm")

sjp.lmer(mod2a, type = "pred", var = c("Session"), show.scatter = FALSE, show.ci = TRUE)
sjp.int(mod2b, swap.pred = T, mdrt.values = "meansd")
```

# 4. Calculate confidence intervals around your estimates for your final models
```{r}
#Confidence intervals around nominal model
confint(mod1a, level = .95)

mod1a.ci <- confint(mod1a, level = .95, oldNames = F, method = "boot", nsim = 100)
broom::tidy(mod1a.ci)

#Confidence intervals around continuous model
mod2a.ci <- confint.merMod(mod2a, level = .95, oldNames = F, method = "boot", nsim = 100)
broom::tidy(mod2a)
```

# 5. Include both types of covariates in a single model. How does your interpretation of parameters change?
```{r}
mod3 <- lmer(Utterances.with.Letters ~ Session*Child.Gender*SES + (Session|Subject), data = children)
summary(mod3)
```
**Fixed Effects** 

Intercept: -0.65; mean of number of utterances with letters for males with mean level SES at Session 0 (around age 10 months)

Session: 1.08; increase in number of utterances with letters every 4 months, for males and when SES at mean level 

Child.GenderFemale: 1.83; the in the number of utterances with letters between males and females when SES at mean level

SES: -1.82; decrease, for males, in number of utterances for every 1 unit increase in SES

Session:Child.GenderFemale: -0.53; the difference between the slopes for males and females when SES at mean level

Session:SES: 0.66; the change in the slope of Session, for males, for every 1 unit increase in SES 

Child.GenderFemale:SES: 2.81; the difference between the relationship of SES and the number of utterances with letters for males and females at Session 0

Session:Child.GenderFemale:SES: -0.77; the difference between the interaction of SES and session for males and females