---
title: "Week 1 Homework"
author: "Molly Farry-Thorn"
date: "September 4, 2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(psych)
library(tidyr)
library(plyr)
```

1. Move your data into a long format and a wide format. Did you have any specific challenges that you encountered? If so, discuss them.

```{r}
letters <- read.delim("Masters coded letters.csv", header=TRUE, sep=",")
long_to_wide <- unite(letters, "Participant", c("Subject","Speaker"))

long_to_wide <- gather(long_to_wide, Var, Val, select=c("Age", "Length", "Utterances", "Utterances.with.Letters", "Letter.name", "Letter.shape", "Letter.sound", "Spelling", "Child.name", "Other.name", "Names", "Letter.Teaching", "Not.Letter.Teaching")) %>%
    unite(VarG, Var, Session) %>%
    spread(VarG, Val)
long_to_wide <- long_to_wide[,c(1:14,18:25,15:17,26:169)]
```

2. Create a wave variable and date variable (if applicable).

I have a variable for the session number and a variable for the child's exact age. I don't have the dates of the sessions

3. What is your sample size for each wave of assessment?
```{r}
ddply(letters, .(Session,Speaker), summarize, N = length(Speaker))


```

4. Take the date variable and convert it to a different date format such as time in study or age (if appropriate). What scale is most suitable for your analyses? (weeks/months/years?)

I don't have the dates of the sessions

5. Graph your data using the different time metrics, fitting individual curves for each person.
```{r}
gg1 <- ggplot(letters,
   aes(Session, Utterances.with.Letters, group = Subject, color = factor(Speaker))) + geom_line() +  geom_point() + facet_wrap( ~ Subject)
gg1

```
6. Create an overall average trend of your data (split up into groups if appropriate). Attempt to color your individual data points and/or shade different lines (highlight some particiapnts, highlight the average trend line but not the individual level lines)
```{r}
gg2 <- ggplot(letters, aes(Session, Utterances.with.Letters, color = factor(Speaker))) + 
       geom_point() + geom_smooth() 
gg2


gg3 <- ggplot(letters, aes(Session, Utterances, color = factor(Speaker))) + 
       geom_point() + geom_smooth() 
gg3
```

7. Look at the correlations of your DV across time
```{r}
func <- function(letters)
{return(data.frame(CORR = cor(letters$Utterances, letters$Utterances.with.Letters)))}
ddply(letters,.(Session), func)
```